\chapter{Theoretical Framework}\label{chap:theory}

In this chapter, I develop the theoretical model that motivates and disciplines my empirical strategy. This framework is not intended to produce direct estimates of structural parameters; rather, it serves to \textbf{derive qualitative predictions} that can be tested through reduced-form political economy analysis. By formalizing the incentive structure of formal-informal linkages, the model makes clear what empirical patterns we should observe under each competing hypothesis, and what we should \textit{not} observe.

Unlike traditional dualist models that view informality as a transitory friction, the framework treats the formal-informal linkage as a structural feature of modern production. By modeling the ``Make vs.\,Buy'' decision through a lens of bargaining and power, I derive testable predictions about how productivity gains, institutional enforcement, and Global Value Chain (GVC) logics shape the \textit{distribution} of economic surplus. These predictions form the benchmark for adjudicating between the \textit{displacement} and \textit{adverse incorporation} hypotheses.

\section{Model Specification}
The central theoretical challenge is to distinguish between two competing explanations for persistent informal sector wage suppression. The \textbf{displacement view} attributes wage gaps to technological backwardness: informal firms produce less value per worker due to limited scale and obsolete equipment, and the gap closes naturally as formal productivity rises. The \textbf{adverse incorporation view} attributes wage suppression to structural power asymmetries: formal buyers exploit monopsony positions to appropriate productivity gains while suppressing the prices paid to informal suppliers.

To derive testable predictions from these competing views, I develop a conceptual model with four components: the firm's decision to outsource, the bargaining game over task prices, the Monopsony-GVC mechanism, and the industrial density effect.

\subsection{The Decision to Outsource}

A representative Formal Firm ($F$) in state $s$ and industry ii
$i$ maximizes profit by choosing between producing tasks in-house (using regulated labor $L_F$) or outsourcing them (using informal job-work $M_{\text{inf}}$).


The production function is modeled as a nested \textbf{Constant Elasticity of Substitution (CES)} technology. For tractability, I impose Constant Returns to Scale (CRS), consistent with the assumption of perfect competition in the product market:

\begin{equation}
	Y_{F,dsi} = A_{F,dsi} \left[ \gamma L_{F,dsi}^\rho + (1-\gamma) M_{\text{inf},dsi}^\rho \right]^{1/\rho}
\end{equation}

where:
\begin{itemize}
	\item $Y_F$ = Output of the formal firm
	\item $A_F$ = Formal firm Total Factor Productivity (Hicks-neutral technical change)
	\item $\gamma \in (0,1)$ = Distribution parameter (share of formal labor)
	\item $\rho < 1$ = Substitution parameter. The elasticity of substitution is $\sigma = \frac{1}{1-\rho}$.
	\item $M_{\text{inf}}$ = Quantity of outsourced informal tasks
\end{itemize}

In textiles, tasks like embroidery, dyeing, or stitching can be moved between a factory floor and outside workshops. The parameter $\rho$ captures this substitutability. When $\sigma$ is high (close substitutes), formal firms can easily switch between in-house production and outsourcing in response to relative price changes. When $\sigma$ is low (complements), the production structure is more rigid.

\subsection{The Institutional Bargaining Game}

Unlike formal labor, which faces statutory minimum wages, the price of outsourced tasks ($p_{\text{sub}}$) is determined through negotiation between formal buyers and informal suppliers. Following the generalized Nash bargaining framework, the equilibrium price is:

\begin{equation}
	p_{\text{sub}} = (1 - \mu_s) \cdot MC_{\text{inf}} + \mu_s \cdot MRP_F
\end{equation}

where:
\begin{itemize}
	\item $MC_{\text{inf}}$ = Marginal cost of informal production
	\item $MRP_F$ = Marginal revenue product of the task for the formal firm
	\item $\mu_s \in [0,1]$ = Informal sector's bargaining power in state $s$
\end{itemize}

When $\mu_s = 0$, formal firms capture all surplus (full monopsony); when $\mu_s = 1$, informal suppliers capture all surplus (competitive outcome).

\subsection{Monopsony and GVC Logic}

A critical addition to this framework is the explicit incorporation of Global Value Chain (GVC) logic. As posited by \textcite{nadvi2004global} and \textcite{chen2007rethinking}, the structure of contemporary manufacturing is defined by "buyer-driven" chains where powerful lead firms capture the majority of rents. In this context, informality is not just a separate sector but a "satellite" system that allows formal lead firms to externalize risks and push costs downward.

Specifically, the formal firm exercises \textbf{monopsony wage-setting power}. Because informal suppliers often lack alternative markets (high relationship-specificity), the formal firm can set a price ($p_{\text{sub}}$) that minimally covers the marginal cost of production without sharing any of the productivity gains ($A_F$). This yields the \textbf{monopsony rent-capture hypothesis}: if extraction dominates, the elasticity of informal wages with respect to formal TFP should be statistically indistinguishable from zero ($E_s \times \ln A_F \to 0$ in low-enforcement environments).

Furthermore, the GVC logic suggests a strategic \textbf{cost-shifting mechanism}. Lead firms increase the volume of tasks outsourced ($Q_{\text{out}}$) to manage production growth while simultaneously squeezing the prices (wages) paid to those suppliers to maintain margins. This predicts that higher outsourcing intensity may be negatively correlated with informal wages, independent of productivity—a prediction I test via mediation analysis.

% THE FLOWCHART (INSERTED HERE)
\begin{figure}[h]
	\centering
	\begin{tikzpicture}[node distance=1.5cm, auto]
		\tikzstyle{block} = [rectangle, draw, fill=blue!10, text width=3cm, text centered, rounded corners, minimum height=1cm]
		\tikzstyle{line} = [draw, -latex', thick]
		\tikzstyle{inst} = [ellipse, draw, fill=red!10, text width=2.5cm, text centered]
		
		\node [block] (formal) {Formal Firm ($A_F$) \\ (ASI)};
		\node [block, below of=formal, yshift=-1.2cm] (decision) {Production Choice};
		\node [block, below left of=decision, xshift=-2.5cm, yshift=-1.5cm] (make) {\textbf{Make} \\ (Regulated Labor) \\ Cost: $w_F$};
		\node [block, below right of=decision, xshift=2.5cm, yshift=-1.5cm] (buy) {\textbf{Buy} \\ (Informal Task) \\ Cost: $p_{sub}$};
		
		\node [inst, right of=decision, xshift= 6cm] (state) {\textbf{The State ($E$)} \\ (Inspections, Laws)};
		
		\node [block, below of=buy, yshift=-1.5cm] (bargain) {\textbf{Bargaining Game} \\ Power $\mu(E)$};
		
		\path [line] (formal) -- (decision);
		\path [line] (decision) -- (make);
		\path [line] (decision) -- (buy);
		\path [line] (buy) -- (bargain);
		\path [line, dashed] (state) |- node[near end, above] {Determines} (bargain);
		
		\node [right of=bargain, node distance=7cm, text width=5cm, align=left] (note) {\footnotesize \textit{Hypothesis: Lax Enforcement ($E \downarrow$) $\to$ Formal Power ($\mu \downarrow$)}};
		\draw [dashed] (bargain) -- (note);
	\end{tikzpicture}
	\caption{The Political Economy of Value Chains}
	\label{fig:model}
\end{figure}

\subsection{Industrial Density and Structural Informality}

A final component of this framework addresses the \textit{scale} of the informal sector. While the displacement hypothesis predicts that formal growth will absorb informal workers, the \textbf{adverse incorporation hypothesis} suggests that productive formal industries actively create "informal satellites" to manage increased output without increasing regulated labor costs. 

I formalize this through the \textbf{Industrial Density Effect}. Let $N_{\text{inf}}$ be the number of informal satellite enterprises supporting a formal industry. In an adverse incorporation regime, $N_{\text{inf}}$ is a function of formal productivity $A_F$:

\begin{equation}
	N_{\text{inf},dsi} = f(A_{F,dsi}, L_{F,dsi}) \quad \text{where} \quad \frac{\partial N_{\text{inf}}}{\partial A_F} > 0
\end{equation}

This prediction is unique to the structuralist view: as formal firms become more productive, they multiply the "density" of the informal sector via subcontracting layers, rather than replacing it.

\subsection{How Enforcement Affects Bargaining Power}

\textbf{The key innovation} of this model is to endogenize $\mu_s$ as a function of the state's enforcement environment ($E_s$).

In states with strict labor law enforcement, informal workers have credible access to formal sector employment at legally mandated wages. This raises their reservation wage---the minimum they will accept for informal work. Conversely, in states with weak enforcement, formal firms routinely violate minimum wage laws, so informal workers cannot credibly threaten to switch to formal employment. Their outside option is weaker, reducing their bargaining power.

Formally, let $R_s$ denote the reservation wage in state $s$. This depends on:
\begin{itemize}
	\item The statutory minimum wage $w^{\text{min}}_s$
	\item The probability of enforcement $\pi(E_s)$, where $\pi'(E_s) > 0$
\end{itemize}

The effective outside option is:
\begin{equation}
	R_s = \pi(E_s) \cdot w^{\text{min}}_s + [1 - \pi(E_s)] \cdot w^{\text{outside}}
\end{equation}

where $w^{\text{outside}}$ is the fallback wage in the absence of enforcement (e.g., subsistence agriculture, casual labor).

In Nash bargaining, the informal surplus is $(p_{\text{sub}} - R_s)$ rather than $(p_{\text{sub}} - MC_{\text{inf}})$. Higher enforcement raises $R_s$, which increases the informal sector's bargaining power:

\begin{equation}
	\mu_s = f(R_s(E_s)) \quad \text{where} \quad \frac{d\mu_s}{dE_s} = f'(R_s) \cdot \pi'(E_s) > 0
\end{equation}

For empirical implementation, I approximate this with a linear reduced form:

\begin{equation}
	\mu_s = \beta_0 + \beta_1 E_s + \epsilon_s
\end{equation}

\textbf{Hypothesis:} $\beta_1 > 0$. States with stricter enforcement exhibit higher informal bargaining power, manifesting as greater ``pass-through'' of formal sector productivity gains to informal wages.

\subsection{From Theory to Reduced-Form Predictions}
\label{subsec:predictions}

The model above is not directly estimated as a structural system --- the substitution parameter $\rho$ and the bargaining power $\mu_s$ are not identified as point estimates in the empirical analysis. This approach aligns with a robust tradition in labour economics of using micro-founded models to \textit{motivate} and \textit{discipline} reduced-form regressions rather than calibrating them via structural methods \autocite{Angrist2010}.

The value of the framework lies in what it implies about the \textit{direction and pattern} of observable reduced-form coefficients. Table~\ref{tab:predictions} summarizes the core predictions:

\begin{table}[H]
\centering
\caption{Theoretical Predictions under Competing Hypotheses}
\label{tab:predictions}
\small
\begin{tabular}{llcc}
\toprule
Mechanism & Empirical Test & Displacement Prediction & Adverse Incorporation Prediction \\
\midrule
Outsourcing elasticity & $\frac{\partial \ln Q_{out}}{\partial \ln A_F}$ & $< 0$ (substitution) & $> 0$ (regulatory arbitrage) \\
Wage pass-through & $\frac{\partial \ln w_{inf}}{\partial \ln A_F}$ & $> 0$ (shared gains) & $\approx 0$ (monopsony) \\
Enforcement interaction & $\frac{\partial w_{inf}}{\partial (E_s \times A_F)}$ & Flat & $> 0$ (enforcement enables sharing) \\
Cost-shifting & $\frac{\partial w_{inf}}{\partial Q_{out}}$ & $0$ & $< 0$ (volume suppresses price) \\
Structural persistence & $\beta_1$ in AR(1) & $\to 0$ (convergence) & $\approx 1$ (lock-in) \\
\bottomrule
\end{tabular}
\end{table}

The key insight is that \textbf{the sign and significance of these reduced-form coefficients is itself the test of the theoretical mechanism}. A positive outsourcing elasticity implies that $\sigma > 1$ (substitution) and that enforcement pressure ($E_s$) raises relative informal task costs. A null wage pass-through coefficient implies that $\mu_s \approx 0$ at the margin: formal firms capture productivity gains because informal suppliers have no credible outside option to negotiate a higher share. I do not need to calculate $\mu_s = 0.05$ to make this inference; observing $\beta_{wages} \approx 0$ \textit{across all enforcement environments} is precisely what $\mu_s \approx 0$ would imply.

This strategy has two advantages over structural calibration. First, it is robust to model misspecification in the CES production function: even if the exact functional form is wrong, the qualitative predictions about signs and interactions remain valid under more general substitution structures. Second, it allows me to test each mechanism separately using the best-suited identification strategy for each prediction, rather than jointly estimating a system that relies on strong functional form assumptions throughout.

The four core empirical tests --- Strategic Outsourcing, Wage Pass-Through, Persistence, and Cost-Shifting --- are thus the \textit{reduced-form equivalent} of the structural tests this model motivates. Each test maps directly onto a parameter or mechanism in the theoretical framework, as detailed in Chapter~\ref{chap:methodology}.
