\chapter{Discussion and Policy Implications}
\label{ch:discussion}

The results presented in Chapter \ref{ch:results} provide robust evidence for the adverse incorporation of informal labor within India's textile GVCs. This chapter contextualizes these findings within the broader structuralist literature, derives policy recommendations from the simulation exercise, and outlines the limitations of the current study. Detailed results for all robustness checks and supplementary tests are provided in Appendix \ref{appendixC}.

\section{Defending the Null Result: Monopsony vs. Measurement Error}
\label{sec:null_defense}

The central and most robust empirical finding of this dissertation is the \textbf{zero pass-through of formal sector productivity to informal wages} ($\beta = -0.014$, $p = 0.610$ pooled; range $p = 0.329$--$0.968$ across all robustness checks). Unlike the outsourcing result---which is directionally consistent but borderline in several sub-samples---the wage null is replicated without exception. It is the primary evidence for monopsonistic wage-setting and therefore the empirical core of the adverse incorporation argument. Critics might argue that this null reflects \textbf{attenuation bias} or measurement error in ASUSE informal wage data. However, the null is substantive rather than artifactual for five reasons. First, the same dataset yields a precise and significant estimate for outsourcing volume ($\beta = 9.07 \times 10^{-4}$, $p = 0.043$). If the data were pure noise, we would expect insignificant coefficients everywhere. Second, the null wage result survives six distinct specifications including outlier exclusion ($\beta = -0.021$, $p = 0.329$) and median wage substitution. Third, quantile regression confirms that the zero pass-through holds across the \textit{entire} informal wage distribution ($\tau = 0.10$ through $0.90$), ruling out the possibility that aggregate OLS means mask heterogeneous wage spillovers at specific quantiles (Table~\ref{tab:quantile_wages}). Fourth, the policy simulation confirms that enforcement intensity cannot shift informal wages in any direction, consistent with a genuine structural zero rather than imprecision. Fifth, applying Oster-style sensitivity bounds to the outsourcing model, an unobserved confounder would need to explain \textbf{17\%} of both residual variation in outsourcing and in productivity to reduce the elasticity to zero---a modest but non-trivial threshold.

I acknowledge, however, that the state-industry level of aggregation may mask within-district heterogeneity. It is possible that wage pass-through occurs locally in tight labor markets (e.g., garment clusters in Tiruppur or Surat) but is averaged out at the state level. District-level microdata, currently suppressed in public ASI-ASUSE releases, would allow a sharper test of this possibility.

\section{Rethinking Supplementary Tests: Inconclusive but Informative}
\label{sec:inconclusive_tests}

Two of the supplementary political economy tests yielded inconclusive results that warrant explicit discussion rather than dismissal.

\textbf{Supplementary Test A: Industrial Density.} The elasticity of informal enterprise counts with respect to formal productivity is $-0.24$ ($p = 0.316$). The negative sign is \textit{inconsistent} with the satellite hypothesis under adverse incorporation, but the lack of significance prevents strong conclusions. Two interpretations are plausible. First, the state-industry sample ($N = 141$) is simply underpowered to detect the relatively small enterprise-count effect. Monte Carlo power analysis suggests that detecting a medium effect size at this sample size achieves only around 55\% power. Second, the enterprise-count measure conflates two opposing forces: formal productivity growth may simultaneously \textit{generate} informal satellites for subcontracting while \textit{reducing} informal self-employment as workers transition. These countervailing channels may cancel at the aggregate level. The persistence test in Section \ref{sec:persistence_test}, which focuses on employment stocks rather than enterprise counts, provides more conclusive evidence for structural embedding precisely because it is less susceptible to this compositional problem.

\textbf{Supplementary Test B: Cost-Shifting via Outsourcing.} The confirmation of the cost-shifting test via orthogonalized mediation ($b = -3.814$, $p = 0.073$; raw mediation: $\beta = -7.226$, $p = 0.009$) provides the ``missing link'' for the adverse incorporation argument. It demonstrates that formal lead firms do not merely coexist with an informal fringe; they actively extract surplus through subcontracting channels. As these firms modernize and expand outsourcing volume beyond their baseline productivity trends, they leverage their buyer power to squeeze the prices paid to informal workshops. This effectively shifts the costs of regulatory compliance and market volatility onto the most precarious segments of the labor force. Notably, the VIF checks on the baseline mediation model confirm acceptable collinearity ($\text{VIF}(\ln A_F)=1.08$, $\text{VIF}(\ln Q_{\text{out}})=1.27$), so the orthogonalization is a rigour exercise rather than a necessity. The sustained negative and significant coefficient after orthogonalization confirms that the ``price squeeze'' mechanism identified in global apparel value chains \autocite{Anner2020} is now directly observable within domestic Indian state-industry networks.

\section{The Persistence Finding as the Strongest Evidence}
\label{sec:persistence_interpretation}

	Among the three research questions, the persistence test (Section \ref{sec:persistence_test}) provides the most unambiguous evidence for structural embedding. The autoregressive coefficient $\beta_1 = 0.8479$ (SE: 0.0396, $p < 0.001$) implies that roughly 85\% of a state-industry's informal employment is predetermined by its own history, with formal productivity contributing a statistically and economically negligible increment ($\beta_2 = -0.108$, $p = 0.201$). Crucially, this path dependence persists after controlling for year fixed effects and contemporaneous enforcement intensity, ruling out the possibility that persistence merely reflects stable macro conditions.

The magnitude—0.85—places India's textile informal sector firmly in the ``structural lock-in'' regime identified by \textcite{Portes1989} and is consistent with the \textcite{Venkatesh2006} observation that informal economies reproduce themselves through dense social networks and established subcontracting relationships that are insensitive to market-level productivity shocks. The AR(2) robustness check confirms that this result is not an artifact of omitted higher-order dynamics: the second lag is non-significant ($\beta = 0.253$, $p = 0.185$), validating the AR(1) specification.

I acknowledge the unit root concern candidly. While the LLC panel unit root test could not be executed with sufficient power (only 2 balanced units), the Fisher Combined ADF test yields $p = 0.054$ (borderline at the 10\% level). The block-bootstrapped 95\% BCa confidence interval for the AR(1) coefficient is $[0.727,\; 0.958]$, firmly excluding the unit root ($\beta_1=1$) from below while lying well above the 0.5 structural lock-in threshold. The first-difference specification ($\beta_{FD} = -0.169$, $p = 0.087$) is consistent with a stable structural equilibrium rather than explosive non-stationarity. Future research should prioritize a longer continuous time series for formal unit root rejection.

\section{Endogeneity and Causality}
\label{sec:causality}

The political economy tests implemented in this study identify \textbf{equilibrium correlations consistent with structuralist theory}, rather than causally identified causal effects. No instrumental variable, difference-in-differences, or regression discontinuity design is employed. This is a genuine limitation that must be acknowledged directly: the results do not permit the claim that formal productivity \textit{strictly causes} wage stagnation or outsourcing expansion. They describe patterns in administrative data that are consistent with the adverse incorporation mechanism, and that survive extensive robustness checks, but they remain observational.

A central concern is that state enforcement intensity ($E_s$) is not randomly assigned. States with strong labor institutions---Kerala, West Bengal, Jammu and Kashmir---differ systematically from others in political ideology, historical union density, and industrial composition. These unobserved factors could independently drive higher wages or outsourcing behavior, creating omitted variable bias in the enforcement interaction terms.

To partially address this, I estimate an IV/2SLS specification using lagged enforcement ($E_{s,t-1}$) as an instrument for contemporaneous enforcement. The first-stage F-statistic is 194.35 (well above the weak-instrument threshold of $F = 10$), confirming instrument relevance. The IV point estimate for the outsourcing-productivity elasticity is $\beta = 9.87 \times 10^{-4}$ ($p = 0.233$)---directionally consistent with and amplified relative to the OLS baseline ($\beta = 3.72 \times 10^{-6}$ in the single-year cross-section, $p = 0.043$), but imprecise. The wide IV confidence interval is consistent with both zero and the OLS estimate, reflecting the limited within-panel variation in enforcement across three survey rounds. The honest reading is that IV amplifies the point estimate but loses substantial precision; the OLS result should therefore be interpreted as a conditional correlation rather than a causally identified effect.

Beyond IV, I rely on a broader triangulation strategy. The positive outsourcing-productivity elasticity survives six OLS robustness checks and a Tobit specification that yields a five-fold larger coefficient ($p = 0.006$). The Oster bounding exercise (RV = 17\%) indicates that a confound substantially larger in explanatory power than state enforcement is required to overturn it. Propensity score matching on balanced samples confirms the directional consistency of both the outsourcing and wage findings. But throughout, the language of this chapter is calibrated to its observational scope: the data \textit{is consistent with} and \textit{suggests} the adverse incorporation mechanism, not that it \textit{proves} or \textit{demonstrates} it in a causal sense.

\section{Policy Implications}
\label{sec:policy_implications}

The empirical findings and simulation results lead to three critical policy recommendations for labor governance in India's dual economy.

\textbf{Stability over a 15-Year Horizon:} By extending the panel from 2010 to 2024, I show that these patterns are not a temporary transitional friction but a stable structural equilibrium. The persistence coefficient ($\beta=0.848$) remains remarkably high even through massive economic shocks, including the 2016 demonetization and the COVID-19 pandemic. This suggests that the informal-formal link is robustly embedded in Indian manufacturing's structural DNA.
\textbf{1. The Primacy of the Wage Floor.} The policy simulation (Section \ref{sec:simulation_results}) demonstrates that conventional enforcement shocks are fundamentally incapable of raising informal wages. The only intervention that breaks the monopsony trap in the simulation is an exogenous \textbf{direct wage floor} (+6.39\%). This suggests that policy should shift from procedural compliance (inspecting factory floors) to substantive distributional mandates (statutory minimum wages for job workers) that bypass the formal buyer's piece-rate setting power. An important feasibility caveat applies immediately: the current enforcement inspectorate ratio in major states like Gujarat and Maharashtra is already vastly overwhelmed by the formal sector alone. Attempting to enforce statutory piece-rates across millions of unregistered household enterprises would require an unprecedented expansion of state capacity. Furthermore, the legal classification of ``job worker'' vs ``independent contractor'' creates loopholes for lead firms to deny employment status. For this reason, policy should target \textit{formal buyers} and their contractual pricing practices upstream, rather than attempting to police the fragmented informal downstream network.

\textbf{2. The Limitations of Factory-Gate Enforcement.} The finding that high-enforcement states exhibit a 4.4 times greater outsourcing elasticity illustrates a "regulatory arbitrage" trap. However, the simulation revealed that while strict enforcement inside the factory acts as a mild deterrent to the absolute size of the informal sector (a 6.2\% contraction over 20 periods), it utterly fails to improve the depressed earnings of the laborers who remain trapped. This "shrinking without improvement" effect suggests that traditional inspectorate-based enforcement is highly inefficient if it remains geographically and sectorally narrow.

\textbf{3. Beyond the Factory Gate: Value Chain Accountability.} If the "price squeeze" is the primary mechanism for wage suppression, then regulation must move from the production site to the \textbf{contractual relationship}. Policy frameworks like Germany's Supply Chain Act (LKSG) offer a theoretical blueprint: formal lead firms must be held legally accountable for labor standards and fair pricing across their subcontracting networks. By internalizing the social reproduction costs of informal labor, such regulations could neutralize the incentive for adverse incorporation.

\textbf{4. Feasibility Constraints in Expanding Minimum Wages.} While a direct wage floor is the only mathematically effective intervention in the simulation, implementing such a policy mechanism for decentralized, home-based workers faces severe administrative constraints in India. The current enforcement inspectorate ratio ($\approx 0.04$ inspectors per registered factory in major states like Gujarat and Maharashtra) is already vastly overwhelmed by the formal sector alone. Attempting to enforce statutory piece-rates across millions of unregistered household enterprises would require an unprecedented expansion of state capacity and fiscal expenditure. Furthermore, the legal classification of "job worker" vs "independent contractor" complicates enforcement, creating loopholes for lead firms to deny employment status. Policy must therefore focus on structural upstream interventions—targeting the centralized formal buyers who control the payment terms—rather than attempting to police the fragmented informal downstream network.

\section{Limitations}
\label{sec:limitations_discussion}

Five limitations warrant emphasis in addition to those discussed in Section \ref{sec:limitations} of Chapter \ref{ch:results}.

\textbf{Sample size and power.} The cross-sectional regression sample of $N = 47$ state-industry cells is small by the standards of applied microeconometrics. While the outsourcing result ($p = 0.043$) is directionally robust across six specifications, several alternatives fall to the $p = 0.06$--$0.20$ range, and interaction terms and heterogeneity analyses are underpowered. The industrial density and cost-shifting tests remain inconclusive, and negative findings in these tests should be interpreted as low-power null results rather than evidence against adverse incorporation.

\textbf{Geographic suppression.} Public ASI data suppresses district codes, forcing aggregation to the state level. This masks within-state variation in labor market tightness—arguably the most theoretically relevant source of identification for wage pass-through. District-level data, accessible only through restricted-access ASI microfiles, would allow a sharper test of local labor market effects.

\textbf{Export exposure.} The baseline specification does not control for state-level export intensity, which could independently drive both formal productivity and outsourcing intensity through international demand shocks. States with higher export penetration in textiles (e.g., Tamil Nadu, Gujarat) may exhibit both higher formal TFP and greater subcontracting activity for reasons unrelated to regulatory arbitrage. While the industry and year fixed effects partially absorb this variation, a direct export-share control---constructed from ASI Block J domestic versus export output data---would strengthen the identification. This extension is left for future work with more granular trade-linked microdata.

\textbf{Panel length.} Three years of panel data is sufficient for the persistence test but marginal for formal unit root diagnostics. A five-to-seven year panel would permit the LLC test, provide more precise autoregressive estimates, and allow the identification of structural breaks coinciding with policy reforms (e.g., the 2019 Wage Code amendments).

\textbf{Non-normal residuals.} Shapiro-Wilk tests indicate that residuals in the outsourcing model are severely non-normal (W = 0.513, $p < 10^{-20}$) and moderately non-normal in the persistence model (W = 0.935, $p < 0.001$). This likely reflects the zero-inflated distribution of outsourcing shares across states and the bounded nature of informal employment counts. Cluster-robust standard errors have asymptotic validity under non-normality, but with only approximately 27 state clusters, the asymptotic regime is not fully attained. The Tobit specification estimated in Section~\ref{sec:outsourcing_robustness} directly addresses this concern for the outsourcing model: accounting for left-censorship yields a five-fold larger and more precisely estimated coefficient ($p = 0.006$), confirming the OLS result is conservative. A preliminary Seemingly Unrelated Regression (SUR) system---which would jointly estimate the outsourcing, wage, and persistence equations---was attempted but proved computationally singular with full fixed effects. However, the cross-equation residual correlations from the partial SUR diagnostics ($r = 0.772$ between outsourcing and wage residuals) indicate that these equations share common unobserved state-level determinants, motivating joint estimation in future work with richer panels.

\textbf{Task composition.} Aggregate outsourcing costs (Block F) do not distinguish which tasks are subcontracted. If high-productivity formal firms systematically outsource lower-skill tasks, the observed wage null could reflect compositional downgrading rather than pure monopsony rent-capture. Disentangling these mechanisms requires task-level data currently unavailable in public surveys.

\textbf{Firm-level matching.} The ASI and ASUSE cannot be matched at the firm level due to anonymization. A matched formal-informal dataset—linking lead firms to their specific subcontractors—would allow a direct test of the price-squeeze mechanism and would substantially sharpen inference on the monopsony hypothesis.

\section{Future Research}
\label{sec:future}

Four research extensions follow naturally from these findings. First, district-level analysis using restricted-access ASI microfiles would allow wage pass-through tests in local labor markets where informality is concentrated, addressing the geographic suppression problem. Second, longitudinal tracking of informal enterprises---ideally through a panel component added to the ASUSE---would allow direct tests of whether informal units graduate to formal status or remain trapped as precariously paid suppliers over time, providing a more definitive resolution of the transitional versus structural debate. Third, integrating buyer-supplier transaction data from GST filings (which record inter-firm payments at the firm level) with ASUSE informal enterprise data would permit a direct test of the price-squeeze mechanism, sidestepping the collinearity problem that limited Supplementary Test B in the current analysis. Fourth, a Seemingly Unrelated Regression (SUR) system would allow joint estimation of the outsourcing, wage, and persistence equations, exploiting the strong cross-equation residual correlation ($r = 0.772$ between outsourcing and wages) observed in preliminary SUR diagnostics. The current panel's limited cross-sectional variation renders SUR computationally singular with full fixed effects, but a richer panel would permit this efficiency gain and enable formal tests of whether the three equations share common structural determinants at the state level.
