\chapter{Conclusion}
\label{ch:conclusion}

This dissertation has investigated the political economy of formal-informal linkages in India's textile and apparel manufacturing, adjudicating between the modernization-driven displacement narrative and the structuralist adverse incorporation narrative. By integrating formal sector ASI data with informal sector ASUSE data into a state-industry panel, and triangulating empirical findings with theoretical simulations, the study provides a comprehensive account of how labor institutions shape surplus distribution in contemporary value chains.

\section{Summary of Findings}
\label{sec:summary_findings}

The research yields four principal conclusions:

First, formal sector productivity growth is positively associated with \textbf{strategic outsourcing} to the informal sector, in a direction consistent with regulatory arbitrage. This relationship is not uniform but is highly sensitive to the regulatory environment: states with stricter labor enforcement exhibit a 4.4 times larger outsourcing elasticity. The result is borderline significant in the baseline specification ($p = 0.043$) and drops below conventional thresholds in several sub-samples. It is best understood as \textbf{suggestive evidence} that informality is strategically deployed as a mechanism for regulatory arbitrage, rather than being a transitory byproduct of low productivity; it is not a definitive causal demonstration.

Second, there is a complete \textbf{absence of productivity pass-through} to informal wages. Across six OLS robustness specifications and quantile regression across the entire wage distribution ($\tau = 0.10$ through $0.90$), the elasticity of informal wages with respect to formal productivity remains statistically zero. Propensity score matching on balanced samples confirms the same conclusion. This confirms a monopsonistic wage-setting regime where formal lead firms capture the entire surplus created by productivity gains, leaving informal workers at near-subsistence earnings regardless of the value they generate.

Third, the informal sector exhibits \textbf{strong structural path dependence}. With an autoregressive coefficient of 0.848 (bootstrapped BCa 95\% CI: $[0.727, 0.958]$), informal employment levels are highly predetermined by their own history and are largely insensitive to contemporaneous formal productivity shocks. Fisher ADF diagnostics ($p = 0.054$) are borderline, precluding definitive stationarity rejection; the result is best framed as near-unit-root path dependence consistent with structural lock-in rather than a cleanly stationary AR(1).

Fourth, \textbf{policy simulations validate the adverse incorporation model}. The simulation shows that conventional labor enforcement merely scales the informal sector (industrial density) without improving its terms of incorporation (wages). The critical divergence—where employment responds to enforcement while wages remain flat—is the hallmark of an adverse incorporation equilibrium.

\section{Contributions of the Study}
\label{sec:contributions}

This research makes three key contributions:

\textbf{1. Theoretical Contribution:} It develops a tractable \textbf{CES-Nash bargaining framework} that nests both displacement and adverse incorporation as competing parameter restrictions. By mapping qualitative industrial narratives to estimable parameters, the model provides a rigorous foundation for future structural research on dual economies.

\textbf{2. Empirical Contribution:} It provides \textbf{direct observational evidence} consistent with the ``price squeeze'' mechanism within India's domestic manufacturing linkages. By identifying the zero pass-through of productivity to wages across six OLS robustness specifications, quantile regression across the wage distribution, and propensity score matching, the study moves beyond purely descriptive accounts of the wage gap to characterize the distributional pattern at the heart of informality. The enterprise-count mechanism for adverse incorporation remains contested---a Negative Binomial specification suggests displacement in enterprise counts ($\beta = -0.589$, $p = 0.001$) even as wages remain suppressed---but the core wage and outsourcing findings (the latter confirmed by a Tobit specification with a five-fold larger coefficient) are robust across alternative estimators. This evidence is correlational rather than causally identified, and future work using matched buyer-supplier microdata would provide sharper causal inference.

\textbf{3. Policy Contribution:} It identifies a \textbf{"regulatory arbitrage trap"} in India's labor governance. The finding that enforcement perversely incentivizes outsourcing to the unregulated informal sector suggests that traditional site-based regulation must be replaced by value-chain-wide accountability frameworks.

\section{Directions for Future Research}
\label{sec:future_directions}

Three directions for future research follow naturally from these conclusions. First, the use of \textbf{restricted-access ASI microdata} would allow for district-level analysis, providing sharper identification of local labor market effects and addressing the geographic suppression problem. Second, the integration of \textbf{GST-linked transaction data} could provide a direct firm-to-firm test of the price-squeeze mechanism, sidestepping the aggregation issues of survey data. Third, \textbf{longitudinal tracking of informal enterprises} would allow researchers to track "graduation" versus "lock-in" dynamics over time, providing a definitive temporal resolution to the dualism debate.

In conclusion, the persistence of informality in India's textile sector is not a sign of modernization's failure, but of its strategic success. As formal firms grow more productive, they do not dissolve the informal sector; they deepen its adverse incorporation. The patterns documented here are observational rather than causally identified, yet they are consistent across administrative data sources, robustness specifications, and a 15-year horizon. Addressing the structural embedding of informality requires a fundamental shift in labor policy—from regulating the factory gate to governing the entire value chain.

